\documentclass[10pt]{article}
 
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb, graphicx, multicol, array}


\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\begin{document}
 
\title{CSCI 301, Winter 2018\\Math Exercises \#7}
\author{Jesse Ericksen - W01173602}
\date{Due: February $28^{th} 2018$}
\maketitle
 
\begin{problem}{1}
A = $\{a^ib^j\mid i \geq j \geq 0\}$
\end{problem}
 
\begin{proof}[Proof]

Suppose A is regular. \newline

Let p = pumping length. 

Let s be some string $\in A$ with $|s| \geq p$. \newline

Suppose $s = a^{p-1} b^{p-1}$. 

Therefore, s is in the language and $|s| > p$.\newline

Suppose $x,y,z$ exist. \newline


Since $|x,y| \leq p$ and $|a| \leq p$, we know atleast that $y$ MUST contain only b's.
\newline

Now, the pumping lemma states $xy^iz \in A$. 

Yet, we can see that for $i > 1$, 

\begin{center}
i.e. $xyyz$ , $xyyyz$ $xyyyyz$ ...
\end{center}

the resulting string CANNOT be in the language A because then we would have more b's than a's which 

would not satisfy the condition that $ i \geq j \geq 0 $. Therefore, this equation cannot be pumped and must 

NOT be regular.

\end{proof}

\begin{problem}{2}
 $A = \{a^{4n+5}b^{3n+2}: n\geq 0\}$
\end{problem}


\begin{proof}[Proof] 

Suppose A is regular. \newline

Let p = pumping length.

Let s be some string $\in$ A with $|s| > p$ \newline

Suppose $ s = a^{4p+5}b^{3p+2}$.

Therefore, s is in the language A and $|s| \geq p$. \newline

Suppose $x,y,z$ exist.

Since, $|xy| \leq p$ we know that x and y must be all a's. \newline

The pumping lemma states that for regular languages with $x,y,z$ that $xy^iz$ must also be in that language. 

However, for any $i > 1$ this is not true for $ s \in A $. \newline

For example, consider $xyyz$

\begin{center}
then $s = a^{4p+5+1}b^{3p+2} = a^{4p+6}b^{3p+2} \notin A$
\end{center}

In this case, any extra a's would result in a string that is not in our language, thus this expression cannot 

be pumped and must NOT be a regular language.





\end{proof}

\begin{problem}{3}
A = $\{a^ib^jc^k \mid i = j \mbox{ or } j = k\}$
\end{problem}

\begin{proof}[Proof]
Suppose A is regular. \newline

Lets break this problem up into two cases and evaluate them individually. \newline

Let p = pumping length.

Let s be some string $\in$ A with $|s| \geq p$ \newline

Case 1: [ i = j ]\newline


Suppose $s = a^{p}b^{p}c^k$. 

Therefore, $s$ is in the language and its length is greater than $p$.
\newline

Since, $|xy| \leq p$, we know $x$ and $y$ must consist of all a's. \newline

We know by the pumping lemma for regular languages that $xy^iz \in A$

Consider one example of this, xyyz. \newline

This results in $s = a^{p+1}b^pc^k$. \newline

Now, $|a| \neq |b|$ which does not satisfy the condition for case 1 that $i = j$. We have a contradiction. \newline \newline

Case 2:  [j = k] \newline

Let $s = a^ib^pc^p$. 

Therefore, s is in the language and $|s| \geq p$. \newline

Suppose i = 0. 

Then for $|xy| \leq p, x,y$ must be all b's. \newline

According to the pumping lemma $xy^iz \in A$. Lets consider $xy^2z$. \newline

This results in $s = a^ib^{p+1}c^p$ and now our condition is not met as $j \neq k$.

Furthermore, this contradiction arises for any value > 1. Since y cannot be pumped without such a 

contradiction, we have proven that this language is NOT regular. 














\end{proof}
\end{document}